---
title: "Top 10 safest neighborhoods in Toronto (2014-2019)"
subtitle: "Based on police reports for six major crimes"
author: "Babak Mokri"
thanks: "Code and data are available at: https://github.com/datababak/2178-A1"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Crime rate statistics in Toronto neighbourhoods can be interesting for anyone who is currently living or planning to move to Toronto. In this paper I have compared six types of major crime data including Assault, Auto Theft, Break & Enter, Robbery, Theft Over, and Homicide, happened in 140 neighbourhoods in Toronto during 2014 to 2019. Then I tried to introduce the top ten safest neighborhoods, however, I found this report could be prone to suffer from some systematic biases as neighborhoods are not equaly divided in terms of area and population. Nevertheless, based on  2016 Census Population, I tried to demonstrate the crime rates for 2019 per 100,000 people."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages("opendatatoronto")
# install.packages("tidyverse")
#install.packages("here")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("ggmap")
#install.packages("ggmap")
#install.packages("kableExtra")
#install.packages("palmerpenguins")

library(tidyverse)
library(ggplot2)
library(here)
library(dplyr)
library(ggmap)
library(ggmap)
library(kableExtra)
library(DT)
library(palmerpenguins)

here::here()

```

# Introduction
Information about the neighborhoods in Toronto has always been a top concern for a newcomer parent like me, having two curious early teenagers who prefer not to stay at home with parents, but explore the neighborhood with their own pals. Tough, I personally encourage them find themselves on their own feet, I should always take some provisions on the background like choosing a safer environments for them to live and flourish. Therefore, I have compared the crime rates between different areas in Toronto during 2014-2019 and also checke the trends to see how crime rates have changed during 2018-2019 in Toronto neighborhoods.

Although the conclusions are prone to initiate some biases, I am just providing information to grab the readers' attention to the fact that what numbers are covering for the different crime rates in each Toronto neighborhood. And also how these numbers can change if you impose your own opinions subjectively In addition, different population densities may result in different conclusions as well. Other demographic factors may also raise this subjectivity, prejudice and partiality. 

In this report I will also elaborate more on selecting different types of variables which initiates biases among people, leading to the same issues when they are chosen for the machine learning practices. 



In this report, I used R [@CitingR] and some of its libraries including *Tidyverse* , *torontoopendata* [@OpenDT]. ggplot2 [@citingggplot2], #Latex#, ((download it in a reproducible way using the R package)) opendatatoronto (Gelfand 2020), and *boom*. This report was then compiled using *R markdown*. I acknowledge that I am not concerned with the reasons behind these numbers nor am I pointing to any specific ward to name it dangerous. 








The last paragraph should set out the remainder of the paper.

# Babak
1. Add an abstract. This should be three or four sentences. And then add a descriptive title (Hint: ‘Paper 1’ is not descriptive.)
2. Add a link to your GitHub repo via a footnote.
3. Check that your GitHub repo is well-organized, and add an informative README. (Hint: Comment. Your. Code.). Make sure that you’ve got at least one R script in there, in addition, to your R Markdown file.

# Data
I used Toronto Police Reports data, taken from the Open Data Toronto Portal [@COT], to explore how crime rates vary between different neighborhoods in Toronto. among many existing reports, I chose the one which contains data for six Major Crime Indicators (MCI) for the years 2014 to 2019. This dataset is updated annually and it was last updated at Sept 18, 2020 by its publisher, Toronto Police Services. It is available in SHP, CSV, GEOJSON, GPKG formats and it can be found here in this [LINK}] (https://open.toronto.ca/dataset/neighbourhood-crime-rates/). The attributes of this dataset can be found in [^appendix 1].

# Part I

For the purpose of choosing a safe neighborhood, I summed up the 6 MCI average crime occurrences during 2014-2019 to consider all repoted crimes for each neighborhood. the first top 10 safest neighbourhoods and the last 5, are demonstrated in (Table1 \@ref(fig:table11)). 

```{r}

```
Then I realized that, for the purpose of raising my teenagers, I should be more interested to see where assult would have happened as my kids would be more the victim of this kind of crimes than ... for example. So I gave weighted those two crimes, and found new results in (Table2 \@ref(fig:table22)). 

```{r}

```
I am looking into this problem from my own perspective, however, what is included in this table for the last 5 neighborhoods, is different from what was there in the first table. 

In this example you may notice how weighting has changed the results, leading to nominating the second table's last 5 neighborhoods as the most dangerous one in Toronto.

I would like to draw you attention to how a subjective system of weighting may result in different clues, leading to emerging future biases. And consider that how this may happen in machine learning algorithms where experts hire some external heuristics for their algorithms to shortcut the processes of their machines and thus, to help the algorithms converge faster. This would be absolutely for their positive desired approached, however, like what happened to our results, it may generate other unforeseen biased outcomes.

# Part II

As many other people usually do, I just relied on the existing dataset, regardless of the area and/or the density of population for each neighborhood. Let me put it this way, does a low rate crime in a small area or very less crowded neighborhood mean that it is safer? 

Let's look at the (Figure1 \@ref(fig:figure11)) to see haw the distribution of population based on 2016 Census and the sum of average reported crimes happened between 2014-2019.

Let's divide the average crime rates for each neighborhood to their population and see the results change at (Table3 \@ref(fig:table33)). 


```{r}

```
As you may notice we have got 3 different results for the safest and the most dangerous Toronto neighborhoods.

Considering that these factors impacts the real estate prices in the neighborhoods, I assume that those areas which are systematically labeled as the most dangerous ones in our first table, are the victim to systematic violence which XXX explains in his article .... in This article he covers the systematic violence for the absence of some existing attributes like which we have not divided neighborhoods equally to realize where you can find most criminals among different Toronto neighborhoods.

# Paer III

Toronto is receiving number of immigrants each year. So, we should expect demographic changes in neighborhoods. There is also on variable in this dataset which shows the trend of crimes change between 2018 to 2019. Figure (Figure1 \@ref(fig:table11)) shows this trend for the 15 neighborhoods introduced in Table 1

```{r}

```

If you see a sharp increase/decrease in ... it is because of lack os data.





 1. Index_:  Record Unique Identifier
2. event_unique_id:  Event Unique Identifier
3. occurrencedate:  Date of occurrence
4. reporteddate:  Date occurrence was reported
5. premisetype:  Premise where occurrence took place
6. ucr_code:  URC Code
7. ucr_ext:  URC Code Extension
8. offence:  Offence related to the occurrence
9. reportedyear:  Year occurrence was reported
10. reportedmonth:  Month occurrence was reported
11. reportedday:  Day occurrence was reported
12. reporteddayofyear:  Day of year Occurrence was reported.
13. reporteddayofweek:  Day of week occurrence was reported
14. reportedhour:  Hour occurrence was reported
15. occurrenceyear:  Occurrence year
16. occurrencemonth:  Occurrence month
17. occurrenceday:  Occurrence day
18. occurrencedayofyear:  Occurrence day of year
19. occurrencedayofweek:  Occurrence day of week
20. occurrencehour:  Occurrence hour
21. **MCI:* Major Crime Indicator related to the offence
22. Division:  Division where event occurred
23. Hood_ID:  Neighbourhood Name
24. Neighborhood:  Neighborhood Identificator
25. Long:  Longitude of point extracted after offsetting X and & Coordinates to nearest intersection node
26. Lat:  Latitude of point extracted after offsetting X and & Coordinates to nearest intersection node
27. ObjectId:  Id of the report






Our data is of penguins (Figure \@ref(fig:bills)).

AND WE ARE USING RRRRR

SO CITING R [@CitingR, page(10)]

SO CITING R [@citingggplot2, page(11)]

```{r table1}
library(kableExtra)

dt <- readr::read_csv(here("inputs/data/a3.csv"))
dt <- dt[-c(11:135),]

kbl(dt) %>%
  kable_paper("striped", full_width = F) %>%
  column_spec(1:2 , bold = T) %>%
  row_spec(11:15, bold = T, color = "white", background = "#D7261E")

```









``` {r graph1}

nbr <- dt$raw_data_AVG.Neighbourhood
avg_crm <- dt$total_AVG_crimes

ggplot(data = dt, aes(x = avg_crm, y = nbr)) + 
  geom_bar(stat = 'identity') +
  xlab("Average crimes between 2014 to 2019") +
  ylab("first 5 and last 5 neighbourhoods") + 
  ggtitle("First 5 and last 5 Neighborhoods")

#plot1

```


```{r table2}
library(kableExtra)
kbl(dt) %>%
  kable_paper("striped", full_width = F) %>%
  column_spec(1:2 , bold = T) %>%
  row_spec(6:10, bold = T, color = "white", background = "#D7261E")

```




### Others



```{r bills, fig.cap="Bills of penguinshh", echo = FALSE}







ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

\newpage

Talk more about itjjjjjjjjj.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data that I sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```








Talk way more about ittttttt. 









Talk way more about ittttttt. 






\newpage

# Appendixxxxx {-}

Apendix 1

	Column	Description
1. *_id*:   Unique row identifier for Open Data database
2. *OBJECTID*:   Autogenerated distinct record identifier
3. *Neighbourhood*:   Name of City of Toronto neighbourhood
4. *Hood_ID*:   City of Toronto neighbourhood identifier
5. *Population*:   2016 Census Population 
6. *Assault_2014*:   Count of assaults for 2014
7. *Assault_2015*:   Count of assaults for 2015
8. *Assault_2016*:   Count of assaults for 2016
9. *Assault_2017*:   Count of assaults for 2017
10. *Assault_2018*:   Count of assaults for 2018
11. *Assault_2019*:   Count of assaults for 2019
12. *Assault_AVG*:   Average Assaults from 2014 - 2019
13. *Assault_CHG*:   % Change in assaults from 2018-2019
14. *Assault_Rate_2019*:   Rate of assaults for 2019 per 100,000 population
15. *AutoTheft_2014*:   Count of auto thefts for 2014
16. *AutoTheft_2015*:   Count of auto thefts for 2015
17. *AutoTheft_2016*:   Count of auto thefts for 2016
18. *AutoTheft_2017*:   Count of auto thefts for 2017
19. *AutoTheft_2018*:   Count of auto thefts for 2018
20. *AutoTheft_2019*:   Count of auto thefts for 2019
21. *AutoTheft_AVG*:   Averate auto thefts from 2014-2019
22. *AutoTheft_CHG*:   % Change in auto thefts from 2018-2019
23. *AutoTheft_Rate_2019*:   Rate of auto thefts for 2019 per 100,000 population
24. *BreakandEnter_2014*:   Count of break and enters for 2014
25. *BreakandEnter_2015*:   Count of break and enters for 2015
26. *BreakandEnter_2016*:   Count of break and enters for 2016
27. *BreakandEnter_2017*:   Count of break and enters for 2017
28. *BreakandEnter_2018*:   Count of break and enters for 2018
29. *BreakandEnter_2019*:   Count of break and enters for 2019
30. *BreakandEnter_AVG*:   Averate break and enters from 2014-2019
31. *BreakandEnter_CHG*:   % Change in break and enters from 2018-2019
32. *BreakandEnter_Rate_2019*:   Rate of break and enters for 2019 per 100,000 population
33. *Homicide_2014*:   Count of homicides for 2014
34. *Homicide_2015*:   Count of homicides for 2015
35. *Homicide_2016*:   Count of homicides for 2016
36. *Homicide_2017*:   Count of homicides for 2017
37. *Homicide_2018*:   Count of homicides for 2018
38. *Homicide_2019*:   Count of homicides for 2019
39. *Homicide_AVG*:   Averate homicides from 2014-2019
40. *Homicide_CHG*:   % Change in homicides from 2018-2019
41. *Homicide_Rate_2019*:   Rate of homicides for 2019 per 100,000 population
42. *Robbery_2014*:   Count of robberies for 2014
43. *Robbery_2015*:   Count of robberies for 2015
44. *Robbery_2016*:   Count of robberies for 2016
45. *Robbery_2017*:   Count of robberies for 2017
46. *Robbery_2018*:   Count of robberies for 2018
47. *Robbery_2019*:   Count of robberies for 2019
48. *Robbery_AVG*:   Averate robberies from 2014-2019
49. *Robbery_CHG*:   % Change in robberies from 2018-2019
50. *Robbery_Rate_2019*:   Rate of robberies for 2019 per 100,000 population
51. *TheftOver_2014*:   Count of thefts over for 2014
52. *TheftOver_2015*:   Count of thefts over for 2015
53. *TheftOver_2016*:   Count of thefts over for 2016
54. *TheftOver_2017*:   Count of thefts over for 2017
55. *TheftOver_2018*:   Count of thefts over for 2018
56. *TheftOver_2019*:   Count of thefts over for 2019
57. *TheftOver_AVG*:   Averate thefts over from 2014-2019
58. *TheftOver_CHG*:   % Change in thefts over from 2018-2019
59. *TheftOver_Rate_2019*:   Rate of thefts over for 2019 per 100,000 population
60. *Shape__Area*:   Autogenerated area measurement
61. *Shape__Length*:   Autogenerated length measurement
62. *geometry*


\newpage

# References






