---
title: "Top 10 safest neighborhoods in Toronto during 2014-2019"
subtitle: "Based on police report for six major crimes"
author: "Babak Mokri"
thanks: "Code and data are available at: https://github.com/datababak/2178-A1"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: " Crime rate statistices in different Toronto neighbourhoods can be interesting for anyone who is currently living or plan to move to Toronto. In this paper I have compared the 2014-2019 Crime Data by 140 Neighbourhood in Toronto for Assault, Auto Theft, Break & Enter, Robbery, Theft Over, and Homicide. I also used the 2016 Census Population to demonstrate the crime rates for 2019 per 100,000 people. *#The results show that ..... is improving and ..... the .... neigbourhood is reporting the highest rate of ......#* "
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction
Getting information about the neighborhoods in Toronto has always been a top concern for a newcomer parent like me, having two curious early teenagers who prefer not to stick at home with parents, but explore the neighborhood with their pals. Tough I personally like to let them find themselves on their own feet and leave it always open and welcoming to take my consultations for them, I should always take some provisions in background like choosing a safer environments for them to flourish. To this end, I have prepared this report to briefly compare the crime rates between different areas in Toronto and also check the trends to see how this rate is growing regardless of the reason behind this growth. At the end of this paper I found that

Although the conclusion is prone to initiate some biases, we are just providing information for grab your attention to the fact that what numbers are saying about the different crime rates in each Toronto neighborhood. Nevertheless, the difference in the area and populations may have a direct influence on your own understanding for each neighborhood, where we are not providing any  conclusion to avoid any prejudice or partiality.

For this report, I used data from opendatatoronto portal and further analysis are performed using turned it to desired information for conducted all analysis using R and its libraries including *Tidyverse*, #Latex#, ((download it in a reproducible way using the R package)) opendatatoronto (Gelfand 2020), and *boom*. This report was then compiled using *R markdown*. 

The last paragraph should set out the remainder of the paper.

# Babak
1. Add an abstract. This should be three or four sentences. And then add a descriptive title (Hint: ‘Paper 1’ is not descriptive.)
2. Add a link to your GitHub repo via a footnote.
3. Check that your GitHub repo is well-organized, and add an informative README. (Hint: Comment. Your. Code.). Make sure that you’ve got at least one R script in there, in addition, to your R Markdown file.

# Data
Toronto 

Our data is of penguins (Figure \@ref(fig:bills)).

AND WE ARE USING RRRRR

SO CITING R [@CitingR, page(10)]

SO CITING R [@citingggplot2, page(11)]



```{r bills, fig.cap="Bills of penguinshh", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

\newpage

Talk more about itjjjjjjjjj.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data that I sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```




Talk way more about ittttttt. 


\newpage

# Appendixxxxx {-}
00000


ooooo

\newpage

# Referencesssss



```{r billss, fig.cap="Bills of penguinshh", echo = FALSE}

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```


