---
title: "Top 10 safest neighborhoods in Toronto (2014-2019)"
subtitle: "Based on police reports for six major crimes"
author: "Babak Mokri"
thanks: "Code and data are available at: https://github.com/datababak/2178-A1"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Crime rate statistics in Toronto neighbourhoods can be interesting for current residence of it as well as for those who are planning to move to Toronto. In this paper I have compared six types of major crimes including Assault, Auto Theft, Break & Enter, Robbery, Theft Over, and Homicide, happened in 140 neighbourhoods in Toronto during 2014 to 2019. Firstly, I tried to list top ten and last five safest Toronto neighborhoods, however, it became so controversial as I looked into that from 3 various perspectives. Then, I made an analogy of this contradictions to the biased results from AI algorithm, and finally, how safe neighborhoods can be victimized to the existing systematic violence."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages("opendatatoronto")
# install.packages("tidyverse")
#install.packages("here")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("ggmap")
#install.packages("ggmap")
#install.packages("kableExtra")
#install.packages("palmerpenguins")


library(tidyverse)
library(ggplot2)
library(here)
library(dplyr)
library(ggmap)
library(ggmap)
library(kableExtra)
library(DT)
library(palmerpenguins)
library(bookdown)
library(tidyr)
here::here()

```

# Introduction

Choosing a safe neighborhood in Toronto has always been a top concern for a newcomer parent like me, especially when dealing with two curious and challenging early teenagers who prefer not to stay at home with parents but explore the whole neighborhood with their own pals. Tough, I personally encourage them find themselves on their own feet, I should always take some provisions on the background like choosing a safer environment for them to live and flourish. To choose a safe neighborhood, I have compared the crime rates for different areas in Toronto during 2014-2019.

Although the conclusions are prone to initiate some biases, I am just providing information to grab the readers' attention to the fact that what numbers are covering for the different crime rates in each Toronto neighborhood. And also, how these numbers can change the results if you impose your own subjective opinions. In addition, different population densities may result in different conclusions as well. Other demographic factors may also raise this subjectivity, prejudice and partiality. 

In this report I will also elaborate more on selecting different types of variables which initiates biases among people, leading to the same issues when they are chosen for the machine learning practices. 

For coding, analysis and data visualization, I used R [@citingR] programing language, and some of its libraries including tidyverse [@tidy], opendatatoronto [@ODT], ggplot2[@ggp], and bookdown[@bookd]. This report was then compiled using R markdown[@Rmd]. I acknowledge that I am not concerned with the reasons behind these numbers nor am I pointing to any specific ward to name it dangerous. 

I start to download the required dataset, make some quantitative analysis from different viewpoints, and finally make some analogies between human rationale and what may happen in AI algorithms as well as unanticipated issues with categorizations in a dataset.




# Data

I used Toronto Police Reports data, taken from the Open Data Toronto Portal [@COT], to explore how crime rates vary between different neighborhoods in Toronto. Among many existing datasets, I chose the one which contains data for six Major Crime Indicators (MCI) including Assault, Auto Theft, Break & Enter, Robbery, Theft Over, and Homicide occurred between the years 2014 to 2019. This dataset is updated annually, and the last update was at Sept 18, 2020 by its publisher, Toronto Police Services. This dataset is available in SHP, CSV, GEOJSON, GPKG formats and it can be found in this [LINK](https://open.toronto.ca/dataset/neighbourhood-crime-rates/). 

In this dataset, there are 62 attributes in separate columns, providing data for 140 Toronto neighborhoods in distinct rows. The list of attributes for this dataset is appended to this report.




## Part I

For the purpose of choosing a safe neighborhood for my residence, I considered all reported crimes for each neighborhood, for which I summed up the 6 MCI averages, occurred during 2014-2019. 

(Table1 \@ref(tab:table11)) shows the first top 10 and the last 5 safest neighborhoods in Toronto.




```{r table11, echo = FALSE, message=FALSE}
# Reading data from the .csv file (created at Data_Cleaning.R)
dt1 <- readr::read_csv(here("inputs/data/Crime-AVGs.csv"))

# Keeping the first 10 and the last 5 records
dt1 <- dt1[-c(11:135),]

#knitr::kable(dt1, caption = "top 10 & last 5 safest neighborhoods")

tb1 <- kbl(dt1) %>%
  kable_paper("striped", full_width = F) %>%
  column_spec(1:2 , bold = T) %>%
  row_spec(11:15, bold = T, color = "white", background = "#D7261E")

tb1
```




Then I realized that, crimes like Auto-theft, Break & Enter, and Theft Over are less likely to happen to a teenager compared to Homicide, Robbery and Assault for which my kids would be more vulnerable. So, I decided to give some weights to the 3 latter crimes to make the results more accurate for my case. Based on my own reasoning, I gave 5 weights to Homicide (as this was scarier to the parents as well), 2 to Robbery and Assault, and 3 to Assault (to which kids are very sensitive and usually victimized).
(Table \@ref(fig:table22)) shows the new results after my modifications.




```{r table22, echo = FALSE, message=FALSE}
# Reading data from the .csv file (created at Data_Cleaning.R)
dt2 <- readr::read_csv(here("inputs/data/Crime-wAVGs.csv"))

# Keeping the first 10 and the last 5 records
dt2 <- dt2[-c(11:135),]

#knitr::kable(dt, caption = "top 10 & last 5 safest neighborhoods")

tb2 <- kbl(dt2) %>%
  kable_paper("striped", full_width = F) %>%
  column_spec(1:2 , bold = T) %>%
  row_spec(11:15, bold = T, color = "white", background = "#D7261E")

tb2
```




Having said that I am organizing these neighborhoods for my own study, and probably would recommend it to other parents, you may notice how a subjective weighting system could label some other safe areas as a *not safe* area. This tag may then become a viral term for that neighborhood and hence, initiate an unfavorable prejudice among people. 

Further to the safety, there are other factors such as affordability, school ranks, and the proximity to my relatives which impact my decision making. So, calling the safety factor as my first concern, I would like to have the first 20 safest neighborhoods in a graph as shown in (Figure \@ref(fig:figure11)), and apply the other factors to each subsequent candidate neighborhoods consequently.





``` {r figure11, fig.cap="The top 20 safest Toronto neighborhoods", echo = FALSE, message=FALSE}
dt3 <- readr::read_csv(here("inputs/data/Crime-wAVGs.csv"))
dt3 <- dt3[c(1:20),]
nbr <- dt3$Neighborhood
avg_crm <- dt3$Average_Crimes_w

dt3 <- dt3 %>%
  mutate(nbr = fct_reorder(nbr, -avg_crm)) # order by average crime

ggplot(data = dt3, aes(x = avg_crm, y = nbr)) + 
  geom_bar(stat = 'identity') +
  xlab("Average crimes between 2014 to 2019") +
  ylab("Average crime rates (weighted)") + 
  ggtitle("Average crime rates (weighted) between 2014-2019")

#plot1
```

The above graph provides a visual information representation for the different safety levels at a glance, for which in the first two tables, I had to calculate them manually.

In this story, I would like to draw your attention to how subjectivity it may lead to emerging future biases. This is not much far from what may happen in Machine Learning algorithms. For example, in ML algorithms where experts may hire some external heuristic algorithms to reduce the "worthless" CPU energy consumption, improve productivity, and help the machine algorithms to converge faster, they may accidentally contribute to generating unforeseen biased outcomes, like how our outcomes tagged some neighborhoods as more dangerous ones. As Karen Hao claims "The introduction of bias isn’t always obvious during a model’s construction because you may not realize the downstream impacts of your data and choices until much later. Once you do, it’s hard to retroactively identify where that bias came from and then figure out how to get rid of it" [@Hao]. Likewise, I agree the experts' intention, in our example, was positively tuning the results for their own objectives, however, their producing unwilling biases at the background may remain hidden for years (or forever) as "unknown-unknowns" [@Hao] leading to controversial results and antagonistic disputes.


## Part II

Like other users, I just relied on the existing data, which is generated by a valid source (Toronto Police Services), and it is announcing the crime rates for each neighborhood, regardless of the density of population and/or the size of each individual area. Let me put it in this way; Does a low-rate crime in a small area or very less crowded neighborhood mean that it is safe?

Fortunately, this dataset has provided one column indicating the rate of 6 above-mentioned MCIs for 2019 per 100,000 population. (Table \@ref(tab:table33)) shows the top 10 and last 5 safest Toronto neighborhoods considering the population in each district. 

```{r table33, echo = FALSE, message=FALSE}

# Reading data from the .csv file (created at Data_Cleaning.R)
dt3 <- readr::read_csv(here("inputs/data/Crime-rate-19.csv"))

# Keeping the first 10 and the last 5 records
dt3 <- dt3[-c(11:135),]

tb3 <- kbl(dt3) %>%
  kable_paper("striped", full_width = F) %>%
  column_spec(1:2 , bold = T) %>%
  row_spec(11:15, bold = T, color = "white", background = "#D7261E")

tb3
```

let's look into the results. (Table \@ref(tab:table44)) compares Table 1 and Table 3 side by side; And you may notice, a complete change in the results.

```{r table44,echo = FALSE, message=FALSE}

# Code adapted from https://stackoverflow.com/a/49044933
kable(cbind(dt1,dt3), format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")

```
Considering that safe/dangerous labels directly impact the real estate prices in each neighborhood, in my opinion, for areas which are systematically labeled as the most dangerous ones in our first table, I claim they are victimized to systematic violence which Os Keyes explains in his article "Counting the Countless" [@Os]. in This article he covers the systematic violence for the absence of some existing attributes like which we have not divided neighborhoods equally to realize where you can find most criminals among different Toronto neighborhoods.



\newpage

# Appendix {-}

Apendix 1
List of attributes of the main dataset  (https://open.toronto.ca/dataset/neighbourhood-crime-rates/).


1. *id*:   Unique row identifier for Open Data database
2. *OBJECTID*:   Autogenerated distinct record identifier
3. *Neighbourhood*:   Name of City of Toronto neighbourhood
4. *Hood_ID*:   City of Toronto neighbourhood identifier
5. *Population*:   2016 Census Population 
6. *Assault_2014*:   Count of assaults for 2014
7. *Assault_2015*:   Count of assaults for 2015
8. *Assault_2016*:   Count of assaults for 2016
9. *Assault_2017*:   Count of assaults for 2017
10. *Assault_2018*:   Count of assaults for 2018
11. *Assault_2019*:   Count of assaults for 2019
12. *Assault_AVG*:   Average Assaults from 2014 - 2019
13. *Assault_CHG*:   % Change in assaults from 2018-2019
14. *Assault_Rate_2019*:   Rate of assaults for 2019 per 100,000 population
15. *AutoTheft_2014*:   Count of auto thefts for 2014
16. *AutoTheft_2015*:   Count of auto thefts for 2015
17. *AutoTheft_2016*:   Count of auto thefts for 2016
18. *AutoTheft_2017*:   Count of auto thefts for 2017
19. *AutoTheft_2018*:   Count of auto thefts for 2018
20. *AutoTheft_2019*:   Count of auto thefts for 2019
21. *AutoTheft_AVG*:   Averate auto thefts from 2014-2019
22. *AutoTheft_CHG*:   % Change in auto thefts from 2018-2019
23. *AutoTheft_Rate_2019*:   Rate of auto thefts for 2019 per 100,000 population
24. *BreakandEnter_2014*:   Count of break and enters for 2014
25. *BreakandEnter_2015*:   Count of break and enters for 2015
26. *BreakandEnter_2016*:   Count of break and enters for 2016
27. *BreakandEnter_2017*:   Count of break and enters for 2017
28. *BreakandEnter_2018*:   Count of break and enters for 2018
29. *BreakandEnter_2019*:   Count of break and enters for 2019
30. *BreakandEnter_AVG*:   Averate break and enters from 2014-2019
31. *BreakandEnter_CHG*:   % Change in break and enters from 2018-2019
32. *BreakandEnter_Rate_2019*:   Rate of break and enters for 2019 per 100,000 population
33. *Homicide_2014*:   Count of homicides for 2014
34. *Homicide_2015*:   Count of homicides for 2015
35. *Homicide_2016*:   Count of homicides for 2016
36. *Homicide_2017*:   Count of homicides for 2017
37. *Homicide_2018*:   Count of homicides for 2018
38. *Homicide_2019*:   Count of homicides for 2019
39. *Homicide_AVG*:   Averate homicides from 2014-2019
40. *Homicide_CHG*:   % Change in homicides from 2018-2019
41. *Homicide_Rate_2019*:   Rate of homicides for 2019 per 100,000 population
42. *Robbery_2014*:   Count of robberies for 2014
43. *Robbery_2015*:   Count of robberies for 2015
44. *Robbery_2016*:   Count of robberies for 2016
45. *Robbery_2017*:   Count of robberies for 2017
46. *Robbery_2018*:   Count of robberies for 2018
47. *Robbery_2019*:   Count of robberies for 2019
48. *Robbery_AVG*:   Averate robberies from 2014-2019
49. *Robbery_CHG*:   % Change in robberies from 2018-2019
50. *Robbery_Rate_2019*:   Rate of robberies for 2019 per 100,000 population
51. *TheftOver_2014*:   Count of thefts over for 2014
52. *TheftOver_2015*:   Count of thefts over for 2015
53. *TheftOver_2016*:   Count of thefts over for 2016
54. *TheftOver_2017*:   Count of thefts over for 2017
55. *TheftOver_2018*:   Count of thefts over for 2018
56. *TheftOver_2019*:   Count of thefts over for 2019
57. *TheftOver_AVG*:   Averate thefts over from 2014-2019
58. *TheftOver_CHG*:   % Change in thefts over from 2018-2019
59. *TheftOver_Rate_2019*:   Rate of thefts over for 2019 per 100,000 population
60. *Shape__Area*:   Autogenerated area measurement
61. *Shape__Length*:   Autogenerated length measurement
62. *geometry*


\newpage

# References






